---
title: "Palmer Penguins Initial Analysis"
author: "Brianna Nason"
format: html
editor: visual
execute: 
  keep-md: true
---

## Palmer Penguin Initial Analysis

This is an analysis of the Palmer's Penguin data set.

## Loading Packages and Data sets

Here we will load the tidyverse packages and penguins data.

```{r}
#Load the tidyverse
library(tidyverse)
library(kableExtra)

#Read the penguins_samp1 data file from github
penguins <- read_csv("https://raw.githubusercontent.com/mcduryea/Intro-to-Bioinformatics/main/data/penguins_samp1.csv")

#See the first six rows of the data we've read in to our notebook
penguins %>%
  head(2) %>%
  kable() %>%
  kable_styling(c("striped", "hover"))
```

## About Our Data

The data we are working with is a data set on Penguins, which includes 8 features measured on 44 penguins. The features included are physiological features (like bill length, bill depth, flipper length, body mass, etc) as well as other features like the year that the penguin was observed, the island the penguin was observed on, the sex pf the penguin, and the species of the penguin.

## Interesting Questions to Ask

-   What is the average flipper length? What about for each species?

-   Are there more male or female penguins? What about per island or species?

-   What is the average body mass? What about by island? By species? By sex?

-   What is the ratio of bill length to bill depth for a penguin? What is the overall average of this metric? Does it change by species, sex, or island?

-   Does average body mass change by year?

## Data Manipulation Tools and Strategies

We can look at individual columns in a data set or subsets of columns in a data set. For example, if we are only interested in flipper length and species, we can `select()` those columns.

```{r}
penguins %>%
  select(species, body_mass_g)
```

If we want to `filter()` and only show certain rows, we can do that too.

```{r}
#we can filter by sex (categorical variables)
penguins %>% 
  filter(species == "chinstrap")

#we can also filter by numerical variables
penguins %>%
  filter(body_mass_g >= 6000)

#we can also do both
penguins %>%
  filter(body_mass_g >=6000 | (island == "Torgersen"))
```

## Answering Our Questions

Most of our questions involve summarizing data, and perhaps summarizing over groups. We can summarize data using the `summarize()` function, and group data using `group_by()`.

Let's find the average flipper length.

```{r}
#Overall average flipper length
penguins %>%
  summarize(avg_flipper_length = mean(flipper_length_mm))

#Single Species Average
penguins %>%
  filter(species == "Gentoo") %>%
  summarize(avg_flipper_length = mean(flipper_length_mm))

#Grouped Average
penguins %>%
  group_by(species) %>%
  summarize(avg_flipper_length = mean(flipper_length_mm))
```

How many of each species do we have?

```{r}
penguins %>%
  count(species)
```

How many of each sex are there? What about by island or species?

```{r}
penguins %>%
  count(sex)

penguins %>%
  group_by(species) %>%
  count(sex)
```

We can use `mutate()` to add new columns to our data set.

```{r}
#To make a permanent change, we need to store the results of our manipulations
penguins_with_ratio <- penguins %>%
  mutate(bill_ltd_ratio = bill_length_mm / bill_depth_mm)

#Average Ratio
penguins %>%
  mutate(bill_ltd_ratio = bill_length_mm / bill_depth_mm) %>%
  summarize(mean_bill_ltd_ratio = mean(bill_ltd_ratio),
            median_bill_ltd_ratio = median (bill_ltd_ratio))

#Average Ratio by Group
penguins %>%
  group_by(species) %>%
  mutate(bill_ltd_ratio = bill_length_mm / bill_depth_mm) %>%
  summarize(mean_bill_ltd_ratio = mean(bill_ltd_ratio),
            median_bill_ltd_ratio = median (bill_ltd_ratio))

```

Average body mass by year

```{r}
penguins %>%
  group_by(year) %>%
  summarize(mean_body_mass = mean(body_mass_g))
```

## Data Visualization With `ggplot()`

-   What is the distribution of penguin flipper lengths?

-   What is the distribution of penguin species?

-   Does the distribution of flipper length depend on the species of penguins?

-   Distribution per year

-   Bill length by year

-   correlation between bill length and bill depth

This plot is used to show the distribution of Flipper Length between penguins.

```{r}
penguins %>%
  ggplot() +
  geom_histogram( aes(x = flipper_length_mm),
                  bins = 15,
                  fill = "forestgreen",
                  color = "black") +

labs(title = "Distribution of Flipper Length (mm)",
     y = "", x = "Flipper Length (mm)") +
  geom_vline(aes(xintercept = mean (flipper_length_mm)), lwd = 2, lty = "dashed") +
  geom_vline(aes(xintercept = median (flipper_length_mm)), color = 'purple', lwd = 2, lty = "dotted")
```

We will now look at the distribution of species

```{r}
penguins %>%
  ggplot() +
  geom_bar(mapping = aes(x = species)) + 
  labs(title = "counts of Penguin Species", x = "species", y = "count")
```

Let's make a scatter plot to see if bill length is correlated with bill depth.

```{r}
penguins %>%
  ggplot() + 
  geom_point(aes(x = bill_length_mm, y = bill_depth_mm, color = species)) +
  geom_smooth(aes(x = bill_length_mm, y = bill_depth_mm, color = species), method = "lm")
```

```{r}
penguins %>%
  summarize(avg_bill_length = mean(bill_length_mm))

t.test(penguins$bill_length_mm, alternative = "greater", mu = 45, corf.level = 0.95)
```
